---
title: "Controlled Vocabularies"
format: html
editor: visual
---

```{r setup, include=FALSE}
library(tibble)
library(dplyr)
library(DT)
```

## 📚 DFO Pacific Science Vocabularies

This table lists machine-readable vocabularies curated by the Data Stewardship Unit to support standardization, integration, and automation across salmon-related research and management programs.

```{r}
# Manual index of vocabularies with live GitHub links
vocab_index <- tribble(
  ~id, ~name, ~description, ~theme, ~last_updated, ~viewer_url, ~json, ~ttl, ~jsonld,
  "spsr", "SPSR Vocabulary", "Summary indices for fisheries stock assessment (FSARs)", "Stock Assessment", "2025-04-10",
  "https://dfo-pacific-science.github.io/data-stewardship-unit/data-standards/spsr-vocab-browser.html",
  "https://raw.githubusercontent.com/dfo-pacific-science/data-stewardship-unit/refs/heads/main/data-standards/vocab/spsr_vocab.json",
  "https://raw.githubusercontent.com/dfo-pacific-science/data-stewardship-unit/refs/heads/main/data-standards/vocab/spsr_vocab.ttl",
  "https://raw.githubusercontent.com/dfo-pacific-science/data-stewardship-unit/refs/heads/main/data-standards/vocab/spsr_vocab.jsonld"
)

vocab_index <- vocab_index %>%
  mutate(
    browser = paste0("[View](", viewer_url, ")"),
    json_link = paste0("[JSON](", json, ")"),
    ttl_link = paste0("[TTL](", ttl, ")"),
    jsonld_link = paste0("[JSON-LD](", jsonld, ")")
  )

datatable(
  vocab_index %>%
    select(name, theme, description, last_updated, browser, json_link, ttl_link, jsonld_link),
  options = list(pageLength = 10, scrollX = TRUE),
  rownames = FALSE,
  escape = FALSE,
  colnames = c("Name", "Theme", "Description", "Last Updated", "Viewer", "JSON", "TTL", "JSON-LD")
)
```

------------------------------------------------------------------------

## 🛠️ TODOs for Improvement

-   [ ] Load vocab metadata from CSV or YAML instead of inline code
-   [ ] Add filters by theme or program
-   [ ] Include persistent w3id URIs once registered
-   [ ] Add stats (e.g., number of terms, match rate with SPSR)
-   [ ] Group vocabularies by domain (e.g., Stock Assessment, Monitoring, Habitat)