---
title: "DFO Salmon Data Controlled Vocabulary"
format: html
---

```{r setup, include=FALSE}
library(readr)
library(tibble)
library(dplyr)
library(DT)
```

## Background

A growing challenge in salmon research, management, and data stewardship is the proliferation of terms, acronyms, and definitions that can vary across documents, programs, and even within the same organization. This can lead to confusion, miscommunication, and inefficiencies in data analysis and reporting.

To address this, the Data Stewardship Unit (DSU) has initiated a **DFO Salmon Data Controlled Vocabulary**. This is a comprehensive, standardized list of terms and definitions related to salmon data and policy used within Fisheries and Oceans Canada (DFO) Pacific Region.

The table below is meant to evolve into an authoritative, community-curated list of salmon data and policy terms used throughout DFO Pacific. It draws on definitions from regulatory documents, scientific literature, internal guidelines, and other sources—consolidating the ones critical for scienctific clarity and data analysis into a single, searchable, filterable reference.

However, this is not a simple list of terms or glossary, under the hood it follows protocols established by the World Wide Web Consortium (W3C) for controlled vocabularies, specifically the [Simple Knowledge Organization System (SKOS)](https://www.w3.org/2004/02/skos/) standard. This allows us to share the vocabulary in a machine-readable format that can be easily integrated into data systems, APIs, and web applications.


## 🗂️DFO Salmon -- Controlled Vocabulary Terms (🚧Under Construction🚧)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load the vocabulary data from the .csv file
vocab <- read_csv("vocab/salmon_terms.csv")



# Create a datatable with search and filter capabilities ordered by term column
vocab <- vocab %>%
  arrange(Term) %>% # Sort by term for better usability
  mutate(Term = paste0("<b>", Term, "</b>")) # Bold the term for emphasis
datatable(
  vocab,
  options = list(pageLength = 15, scrollX = TRUE),
  rownames = FALSE,
  filter = "top",
  escape = FALSE,
) |> 
  formatStyle(columns = c(1:ncol(vocab)), # Apply to all columns
              whiteSpace = "nowrap",
              overflow = "hidden",
              textOverflow = "ellipsis"
)


```


Under the hood, the vocabulary is formatted in SKOS (Simple Knowledge Organization System), a W3C-recommended framework for representing controlled vocabularies and thesauri. By modeling terms and their relationships (broader, narrower, related), SKOS makes it easy to:

- Maintain consistent labels, definitions, and concept hierarchies.

- Share and integrate the vocabulary with other systems that understand RDF.

The vocabulary is intended to be a living document that evolves with the needs of the community. It is maintained by the Data Stewardship Unit (DSU) within the Fisheries and Assessment Data Section (FADS) in the Pacific Region of Fisheries and Oceans Canada (DFO).

To maximize accessibility, we publish three serializations:

- JSON – A lightweight format ideal for most web and desktop applications that need to load term lists and definitions without RDF tooling.

- JSON-LD – Embeds linked-data context so that terms become addressable URIs; perfect for web pages, APIs, and any environment where you want automatic context resolution and richer interoperability.

- Turtle (.ttl) – A compact, human-readable RDF syntax for semantic-web tools, triple stores, and SPARQL queries, enabling advanced semantic reasoning or integration into broader data ecosystems.

You can download or query any of these formats directly from our GitHub repo for programmatic access, metadata validation, integration, or automated workflows.


File a GitHub Issue at
https://github.com/dfo-pacific-science/data-stewardship-unit/issues

Or contact the Data Stewardship Unit.

<!-- ## 🛠️ TODOs for Improvement

-   [ ] Load vocab metadata from CSV or YAML instead of inline code
-   [ ] Add filters by theme or program
-   [ ] Include persistent w3id URIs once registered
-   [ ] Add stats (e.g., number of terms, match rate with SPSR)
-   [ ] Group vocabularies by domain (e.g., Stock Assessment, Monitoring, Habitat) -->