---
title: "Accessing Data from the Open Government Portal"
format: html
version: "v1.0"
last_updated: "2025-08-11"
---

## Background

This tutorial provides an overview for how to extract data from the [Open Government Portal](https://open.canada.ca/), commonly referred to as *Open Data*.

This tutorial reviews:

-   Querying ArcGIS REST services

-   Extracting data from file geodatabases

-   Using `ckanr` to build more robust code

The examples cover the basics of mapping in R, including plotting both vector and raster data with `ggplot()`, and creating `leaflet` maps. Many online tutorials provide further detail for specifics of mapping in R if you need more background. For example, [this tutorial from the University of Toronto](https://mdl.library.utoronto.ca/technology/tutorials/introduction-gis-using-r) provides a very good overview.

Note that much of this content has been expanded from a [similar tutorial created in the DFO Maritimes Region](https://github.com/AtlanticR/R_Git_Support/blob/main/open_data.Rmd), but with DFO Pacific examples. Other clarifying comments have also been added.

### Loading packages

The [librarian](https://cran.r-project.org/web/packages/librarian/vignettes/intro-to-librarian.html) package is useful for loading multiple packages in R. It allows you to load packages from CRAN, GitHub, and other sources with a single command. It also installs missing packages automatically.

```{r setup-librarian, message=FALSE, warning=FALSE}
# Set CRAN mirror to avoid errors
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Install librarian if not already installed
if (!requireNamespace("librarian", quietly = TRUE)) {
  install.packages("librarian")
}

# Load packages using librarian
librarian::shelf(
  arcpullr, ckanr, dplyr, devtools, ggplot2, ggspatial, leaflet, leafpop, mapview,
  rnaturalearth, sf, stars, terra, viridisLite,
  "github::ropensci/rnaturalearthhires" # example for loading from GitHub
)

```

### Preparing a basemap

There are some great packages to get basemaps in R. The `rnaturalearth` package is a good option for getting country borders and other geographical features. You can use the `ne_countries()` function to get country borders.

In British Columbia, `rnaturalearth` may not have the level of detail required (e.g., some islands are missing). The `bcmaps` package has a lot of detailed basemaps and spatial data for the BC coast, and the `bc_bound_hres()` function is particularly useful for mapping coastline.

Because the data used in the tutorial are for the entire BC coast, `rnaturalearth` is used in the examples below.

```{r coastline, warning=FALSE, message=FALSE}

# Get coastline for North America 
coast = ne_countries(scale = 10, returnclass = c("sf"), continent = "North America") 

```

### Make a quick map 

```{r plot-coast, warning=FALSE, message=FALSE}
# Use ggplot to map the data 
ggplot()+
  geom_sf(data=coast, fill="lightgray")+
  theme_minimal()

```

## Accessing Open Data via ARCGIS REST

The `arcpullr` package is a great way to access [ArcGIS REST services](https://geoappext.nrcan.gc.ca/arcgis/sdk/rest/index.html#//02ss00000057000000), which are often used for storing spatial data on Open Data. You can use the `get_spatial_layer()` function to retrieve a spatial layer from an ArcGIS server.

### Example: Pacific Recreational Fishery Salmon Head Depots

There can many URLs associated with each dataset. The "ArcGIS Rest URL" we need to feed into the `get_spatial_layer()` function should be the one that includes "arcgis/rest/services".

For example, with the [Pacific Recreational Fishery Salmon Head dataset](https://open.canada.ca/data/en/dataset/3cc03bbf-d59f-4812-996e-ddc52e0ba99e), the URL we are interested in is <https://egisp.dfo-mpo.gc.ca/arcgis/rest/services/open_data_donnees_ouvertes/pacific_recreational_fishery_salmon_head_depots/MapServer/>.

If you have trouble tracking this down, go to the appropriate [dataset page](https://open.canada.ca/data/en/dataset/3cc03bbf-d59f-4812-996e-ddc52e0ba99e) on Open Data.

Then, under the **Data and Resources** section, find the item labelled **ESRI REST** (there is both an English and French example) and left click on it. On the [new page that opens](https://open.canada.ca/data/en/dataset/3cc03bbf-d59f-4812-996e-ddc52e0ba99e/resource/1c89a180-7622-42b1-9edf-be2457265880), right click on the **Go to Resource** button, and select **Copy link address**. This will copy the URL you need.

It is also important to specify the appropriate layer. In the example below, the "0" at the end of the address in the the example below denotes the English version, while the 1 represents French (see layers within the [MapServer page](https://egisp.dfo-mpo.gc.ca/arcgis/rest/services/open_data_donnees_ouvertes/pacific_recreational_fishery_salmon_head_depots/MapServer)). They also do not always correspond to language. For example, in the [Pacific Commercial Salmon In Season Catch Estimates dataset](https://egisp.dfo-mpo.gc.ca/arcgis/rest/services/open_data_donnees_ouvertes/pacific_commercial_salmon_in_season_catch_estimates_en/MapServer), layers 0 to 3 represent gill nets, troll, seine and Pacific Fishery Management Areas, respectively.

```{r salmonDepots, warning=FALSE}
salmonDepots = get_spatial_layer("https://egisp.dfo-mpo.gc.ca/arcgis/rest/services/open_data_donnees_ouvertes/pacific_recreational_fishery_salmon_head_depots/MapServer/0")
```

It may be useful to crop the plot to the bounding box of the data. The `st_bbox()` function from the `sf` package allows you to get the bounding box of a spatial object. You can use the `xlim` and `ylim` arguments in the `coord_sf()` function to set the limits of the x and y axes.

```{r depot-map, warning=FALSE, message=FALSE}

# Define lat/lon bounds of your dataset
bounds = st_bbox(salmonDepots)

ggplot()+
  geom_sf(data = coast, fill = "lightgrey")+
  geom_sf(data = salmonDepots, size = 3.5, aes(fill = ACCESSIBILITY___ACCESSIBILITÉ), pch = 21, alpha = 0.85)+
  coord_sf(xlim = c(bounds[1], bounds[3]), 
           ylim = c(bounds[2], bounds[4]))+
  # I'm not always convinced maps north arrows need to be added when the lat/lon is specified but some people insist!
  # Add north arrow to the top right (tr) of the map. It points to "true" north, not magnetic north
  annotation_north_arrow(location = "tr", which_north = "true",
                         #height = unit(0.7, "cm"), width = unit(0.7, "cm"),
                         style = north_arrow_minimal)+
  # Also, scale bars shouldn't be added when you're this zoomed out, but here's how to add one anyway to the bottom left (bl):
  annotation_scale(location = "bl", text_cex = 0.8)+
  scale_fill_discrete(name = "Accessibility")+
  theme_bw() # tidies up the visuals 
  
```

## Accessing file geodatabases

[File geodatabases](https://pro.arcgis.com/en/pro-app/latest/help/data/geodatabases/manage-file-gdb/file-geodatabases.htm) are a proprietary file format developed by ESRI for spatial and non-spatial data. Many spatial datasets are uploaded to Open Data as file geodatabases, and it is possible to download and extract them with R!

The example below shows how to access the file geodatabase from the [floating infrastructure dataset on Open Data](https://open.canada.ca/data/en/dataset/049770ef-6cb3-44ee-afc8-5d77d6200a12/resource/c32dff08-973e-4141-86e8-27839636bc63)

### Creating temporary directories

For exploring data, it is often helpful to create temporary directories to store and extract data. This can be easier than downloading the data directly and setting your directories.

```{r tempDirs, warning=FALSE, message=FALSE}

# Create a temporary directory to store the downloaded zip file
temp_dir = tempdir()

# Define the path for the downloaded zip file inside the temp directory
zip_file = file.path(temp_dir, "zipPath.gdb.zip")

# Define the directory where the zip contents will be extracted
# This is a relative path, so files will be extracted into the current working directory
unzip_dir = file.path("extracted_fgdb")

# Download the zipped geodatabase. To get the correct link, right click on the "Go to Resource" button, and select "Copy link address"
download.file(
  "https://api-proxy.edh-cde.dfo-mpo.gc.ca/catalogue/records/049770ef-6cb3-44ee-afc8-5d77d6200a12/attachments/floating_infrastructure-docks.gdb.zip",
  destfile = zip_file,
  mode = "wb"
)

# Create the extraction directory if it doesn't already exist
dir.create(unzip_dir, showWarnings = FALSE)

# Unzip the downloaded file into the extraction directory
unzip(zip_file, exdir = unzip_dir)

# List the files extracted to verify contents
list.files(unzip_dir)

# Load the layers available in the extracted .gdb file
# Copy the name from the list.files() command. This should match the name of the gdb from the URL above
layers = st_layers(file.path(unzip_dir, "floating_infrastructure-docks.gdb"))

# Turn layers from a list into a dataframe
# layers has 1 entry called docks, so now a df called 'docks' will be created
for(l in layers$name){
  message("Reading layer: ", l)
  assign(l, st_read(file.path(unzip_dir, "floating_infrastructure-docks.gdb"), layer = l, quiet = TRUE))
}

```

### Map the data

```{r dock-map, warning=FALSE, message=FALSE}

# Note that the coordinate reference system of these two layers are different
invisible(st_crs(docks)) # NAD 83
invisible(st_crs(coast)) # WGS 84

# To plot with ggplot, one of them has to be reprojected. I've transformed the docks CRS to match the coastline CRS 
docks = st_transform(docks, st_crs(coast))

# Redefine the bounds of the layer of interest
boundsDOcks = st_bbox(docks)

# Make a map of the docks
ggplot()+
  geom_sf(data = coast, fill = "lightgrey")+
  geom_sf(data = docks, size = 3.5, pch = 21, aes(fill = region), alpha = 0.65)+
  coord_sf(
    xlim = c(boundsDOcks["xmin"], boundsDOcks["xmax"]),
    ylim = c(boundsDOcks["ymin"], boundsDOcks["ymax"])
  ) +
  scale_fill_discrete(name = "Region")+
  theme_bw()

```

## Plotting raster data

An example is shown here for accessing and plotting raster data stored in a geodatabase. You can follow the example in the [Maritimes tutorial](https://github.com/AtlanticR/R_Git_Support/blob/main/open_data.Rmd) for accessing raster data from tiff files, if you are interested.

The example I'm using is from the [recreational vessel traffic model on the BC coast](https://open.canada.ca/data/dataset/fed5f00f-7b17-4ac2-95d6-f1a73858dac0).

```{r temp-vessels, echo = TRUE, message= FALSE, warning= FALSE, error= FALSE, cache= FALSE, results='hide'}

# Follow steps above to create a temporary folder to download & extract geodatabase info
temp_dir = tempdir()
zip_file = file.path(temp_dir, "bc_boating_model.zip")
unzip_dir = file.path(temp_dir, "bc_boating_data")

# Download the dataset zip file
download.file(
  url = "https://api-proxy.edh-cde.dfo-mpo.gc.ca/catalogue/records/fed5f00f-7b17-4ac2-95d6-f1a73858dac0/attachments/Recreational_Boating_Data_Model.gdb.zip",
  destfile = zip_file,
  mode = "wb"
)

# Unzip the contents
dir.create(unzip_dir, showWarnings = FALSE)
unzip(zip_file, exdir = unzip_dir)

# List the file names you just downloaded 
boat_file = list.files(unzip_dir, full.names = TRUE)

```

## Mapping the survey effort data with `ggplot`

Unfortunately, raster data in this format cannot be directly mapped with `ggplot`. But, you can convert it to a data frame and keep the coordinate positions.

If you read through [the metadata](https://egisp.dfo-mpo.gc.ca/arcgis/rest/services/open_data_donnees_ouvertes/recreational_vessel_traffic_model_for_british_columbia/MapServer), you'll see that that many types of data are available in this geodatabase (e.g., point data, raster, vector grid). The data being plotted in this example below is the *surveyeffort* raster dataset, which is specified using the `sub` argument in the `rast()` function.

```{r survey-raster, warning=FALSE, message=FALSE}

# Turn the "survey effort" data into raster format
surveyRast = rast(boat_file, sub = "surveyeffort")

# Use the project function to project to WGS 84 (i.e., to match the CRS of the coastline file)
rast_proj = project(surveyRast, coast)

# Convert it to a dataframe. Keep the coordinate info
rast_df = as.data.frame(rast_proj, xy = TRUE) 


# Plot it with geom_tile and add the coast data as well
ggplot()+
  geom_tile(data = rast_df, aes(x=x, y=y, fill = Recreational_Boating_Data_Model)) +
  geom_sf(data = coast, fill = "lightgrey")+
  scale_fill_viridis_c(option = "magma")+
  coord_sf(
    xlim = c(bounds["xmin"], bounds["xmax"]),
    ylim = c(bounds["ymin"], bounds["ymax"])
  )+
  labs(x = NULL, y = NULL)+
  theme_bw()

```

### Making Leaflet Maps

[Leaflet maps](https://rstudio.github.io/leaflet/) are interactive maps that can be a great way to explore data. Leaflet maps can also be great for point data, since you can create [popup labels](https://rstudio.github.io/leaflet/articles/popups.html) where you can click and view your attribute data.

```{r leaflet, warning=FALSE, message=FALSE}

# Convert raster to leaflet-compatible format
pal = colorNumeric(palette = "magma", domain = values(rast_proj), na.color = "transparent")

# Create leaflet map
leaflet() %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>%
  addRasterImage(rast_proj, colors = pal, opacity = 0.7, project = T) %>%
  addLegend(pal = pal, values = values(rast_proj),
            title = "Survey Effort",
            position = "bottomright")

```

## Using `ckanr` to build more robust code

Sometimes, IT needs to upgrade server hardware or software, which can result in changes to the ArcGIS REST service URLs (i.e., the portion like this: *https://egisp.dfo-mpo.gc.ca/arcgis/rest/services*). However, the Universally Unique Identifier (UUID) associated with each dataset remains unchanged.

In the Open Data portal, the UUID is the final part of the URL after "dataset". For example, in the Pacific Recreational Salmon Head Depots dataset *(https://open.canada.ca/data/dataset/3cc03bbf-d59f-4812-996e-ddc52e0ba99e)*, the UUID is *3cc03bbf-d59f-4812-996e-ddc52e0ba99e*.

To make our script more robust, we can use the CKAN API. CKAN is a data management system that supports publishing, sharing, and discovering datasets. The ckanr package in R allows us to interact with CKAN-based portals, such as Open Data. You can use the `ckanr_setup()` function to establish a connection.

Instead of hardcoding the ArcGIS REST URL in our script (which may change) we can use the CKAN API to reference the stable UUID. This allows us to dynamically retrieve the current ArcGIS REST URL associated with that UUID, ensuring our code remains functional even if updates occur.

### Downloading data

We'll use the [Pacific Recreational Fishery Salmon Head Depots](https://open.canada.ca/data/dataset/3cc03bbf-d59f-4812-996e-ddc52e0ba99e) data as an example. We can use the `ckanr` package to extract the ArcGIS REST URL by referencing the UUID. This may seem like an extra step while you're writing your code, but it will save you time in the long run if the URLs change (it can happen!!)

```{r ckanr, warning=FALSE, message=FALSE}

ckanr_setup("https://open.canada.ca/data")
uuid = "3cc03bbf-d59f-4812-996e-ddc52e0ba99e"
pkg = ckanr::package_show(uuid)
resources = pkg$resources
df_resources = as.data.frame(do.call(rbind, resources))

# Now you can grab the correct ESRI Rest URL ("url"). Make sure to search for that format, and also specify you want the English ("En") version and not French
salmon_url = unlist(df_resources[df_resources$format == "ESRI REST" & sapply(df_resources$language, function(x) "en" %in% x), "url"])

```

### Searching for data

You can also use the `ckanr` package to search for data using the `package_search()` function. You can use the `q` argument to search for key words, and the `fq` argument to specify file types.

```{r searching, echo = TRUE, message= FALSE, warning= FALSE, error= FALSE, cache= FALSE, results='hide'}

ckanr_setup("https://open.canada.ca/data")
search_results = package_search(q = "Pacific salmon")
search_results # Note that UUID also shows up here after the <CKAN Package> text

# You  can also search specifically for spatial data (e.g., shapefiles, geodatabases, etc.) by specifying the fq argument
search_results = package_search(
  q = "Pacific salmon",
  fq = 'res_format:(SHP OR GDB OR GeoJSON OR KML OR CSV)' # CSVs sometimes contain spatial data, sometimes not
)
```
