---
title: "Accessing Data from the Open Government Portal"
format: html
version: "v1.0"
eval: false
last_updated: "2025-08-11"
---

## Some sort of background
This tutorial provides an overview for how to extract data from the [Open Government Portal](https://open.canada.ca/), 
commonly known as Open Data. 

Much of the content has been adapted from a [similar tutorial created in the DFO Maritimes Region](https://github.com/AtlanticR/R_Git_Support/blob/main/open_data.Rmd), but has been adapted for Pacific code.

A few examples have been changed or expanded upon to reflect differences in data.

This tutorial reviews:

* Basics of mapping in R
* Querying ArcGIS REST with arcpullr function






There are many





## Loading packages

I've been using [librarian](https://cran.r-project.org/web/packages/librarian/vignettes/intro-to-librarian.html) to load packages in R. It allows you to load packages from CRAN, GitHub, and other sources with a single command. It also installs missing packages automatically.

```{r, eval=FALSE}
install.packages("librarian")
```

If you already have the librarian package, skip to loading the packages:

```{r,warning=FALSE,message=FALSE}
librarian::shelf(
                 arcpullr,
                 ckanr,
                 dplyr,
                 devtools,
                 ggplot2,
                 ggspatial,
                 leaflet,
                 leafpop,
                 mapview,
                 rnaturalearth,
                 sf,
                 stars,
                 terra,
                 viridisLite)

# This one's always under development though
devtools::install_github("ropensci/rnaturalearthhires")

```


## Preparing a basemap

### Using rnaturalearth for a basemap

There are some great packages to get basemaps in R. The `rnaturalearth` package is a good option for getting country borders and other geographical features. You can use the `ne_countries()` function to get country borders.


```{r coast}
################################################################################


# Get coastline for North America 
coast = ne_countries(scale = 10, returnclass = c("sf"), continent = "North America") 


```

# Make a quick map to see what it looks like
```{r plot coast, message=FALSE}
ggplot()+
  geom_sf(data=coast, fill="lightgray")+
  theme_minimal()


```

## Accessing Open Data

### Querying ArcGIS REST with arcpullr

The `arcpullr` package is a great way to access ArcGIS REST services. It allows you to query and download data from ArcGIS servers. You can use the `get_spatial_layer()` function to get a spatial layer from an ArcGIS server.


Pacific Recreational Fishery Salmon Head Depots on [Open Data](https://open.canada.ca/data/en/dataset/3cc03bbf-d59f-4812-996e-ddc52e0ba99e).

The following code will download and load the entire dataset into R.

```{r draft sites,warning=FALSE}
# Download some data from open data
# For all datasets from Open Data, look carefully at the "Layers" option. For this, 0 means English, and 1 means French
# This is the final thing specified in the URL
salmonDepots = get_spatial_layer("https://egisp.dfo-mpo.gc.ca/arcgis/rest/services/open_data_donnees_ouvertes/pacific_recreational_fishery_salmon_head_depots/MapServer/0")
```


I'm also 'cropping' the plot to the bounding box of the data. The `st_bbox()` function from the `sf` package allows you to get the bounding box of a spatial object. You can use the `xlim` and `ylim` arguments in the `coord_sf()` function to set the limits of the x and y axes.

```{r draft sites plot}
# Define lat/lon bounds, otherwise the map will show the entire globe (see map above)
bounds = st_bbox(salmonDepots)

# Create a map of this data using ggplot
ggplot()+
  geom_sf(data = coast, fill = "lightgrey")+
  geom_sf(data = salmonDepots, size = 3.5, aes(fill = ACCESSIBILITY___ACCESSIBILITÉ), pch = 21, alpha = 0.85)+
  coord_sf(xlim = c(bounds[1], bounds[3]), 
           ylim = c(bounds[2], bounds[4]))+
  # I'm not always convinced maps north arrows need to be added when the lat/lon is specified but some people insist!
  # Add north arrow to the top right (tr) of the map. It points to "true north", not magnetic north
  annotation_north_arrow(location = "tr", which_north = "true",
                         #height = unit(0.7, "cm"), width = unit(0.7, "cm"),
                         style = north_arrow_minimal)+
  # Also, scale bars shouldn't be added when you're this zoomed out, but here's how to add one anyone:
  annotation_scale(location = "bl", text_cex = 0.8)+
  scale_fill_discrete(name = "Accessibility")+
  theme_bw() # tidies up map colouring
  
```

## Downloading geodatabases

# Geodatabases are an ESRI thing, but it is still possible to download and extract the data with R!

# I couldn't find any good Pacific salmon examples so I found one of "floating infrastructure"
# https://open.canada.ca/data/en/dataset/049770ef-6cb3-44ee-afc8-5d77d6200a12/resource/c32dff08-973e-4141-86e8-27839636bc63

``` {r, echo = TRUE, message= FALSE, warning= FALSE, error= FALSE, cache= FALSE, results='hide'}

# Instead of downloading the data directly to your computer, you can create a temporary directory and download it there
# Create a temporary directory to store the downloaded zip file
temp_dir = tempdir()

# Define the path for the downloaded zip file inside the temp directory
zip_file = file.path(temp_dir, "zipPath.gdb.zip")

# Define the directory where the zip contents will be extracted
# This is a relative path, so files will be extracted into the current working directory
unzip_dir = file.path("extracted_fgdb")

# Download the zipped geodatabase. To get the correct link to download, right click on the "Go to Resource" button, and select "Copy link address"
download.file(
  "https://api-proxy.edh-cde.dfo-mpo.gc.ca/catalogue/records/049770ef-6cb3-44ee-afc8-5d77d6200a12/attachments/floating_infrastructure-docks.gdb.zip",
  destfile = zip_file,
  mode = "wb"
)

# Create the extraction directory if it doesn't already exist
dir.create(unzip_dir, showWarnings = FALSE)

# Unzip the downloaded file into the extraction directory
unzip(zip_file, exdir = unzip_dir)

# List the files extracted to verify contents
list.files(unzip_dir)

# Load the layers available in the extracted .gdb file
# Copy the name from the list.files() command. This should match the name of the gdb from the URL above
layers = st_layers(file.path(unzip_dir, "floating_infrastructure-docks.gdb"))

# Turn layers from a list into a dataframe
# layers has 1 entry called docks, so now a df called 'docks' will be created
for(l in layers$name){
  message("Reading layer: ", l)
  assign(l, st_read(file.path(unzip_dir, "floating_infrastructure-docks.gdb"), layer = l, quiet = TRUE))
}

```

### Map the data 

```{r}
### Create a map to show off your work

# Note that the coordinate reference system of these two are different
st_crs(docks) # Nad 83
st_crs(coast) # WGS 84

# To plot with ggplot, one of them has to be reprojected. I've transform the docks CRS to match the coast CRS 
docks = st_transform(docks, st_crs(coast))

# Redefine the bounds of the layer of interest
bounds = st_bbox(docks)

# Make a map of the docks
ggplot()+
  geom_sf(data = coast, fill = "lightgrey")+
  geom_sf(data = docks, size = 3.5, pch = 21, aes(fill = region), alpha = 0.65)+
  coord_sf(
    xlim = c(bounds["xmin"], bounds["xmax"]),
    ylim = c(bounds["ymin"], bounds["ymax"])
  ) +
  scale_fill_discrete(name = "Region")+
  theme_bw()

```

## Plotting raster data 

This is actually raster data stored in a geodatabase. Which is a bit different to plot.
Remi's example shows how to plot tiff data. This is different

If you read through the data, you'll see that that many types of data are availble for this (e.g., point data, raster, vector grid) 
The raster data I'm interested in is titled "surveyeffort" within the  https://egisp.dfo-mpo.gc.ca/arcgis/rest/services/open_data_donnees_ouvertes/recreational_vessel_traffic_model_for_british_columbia/MapServer

``` {r, echo = TRUE, message= FALSE, warning= FALSE, error= FALSE, cache= FALSE, results='hide'}

# Follow steps above to create a temporary folder to download & extract geodatabase info
temp_dir = tempdir()
zip_file = file.path(temp_dir, "bc_boating_model.zip")
unzip_dir = file.path(temp_dir, "bc_boating_data")

# Download the dataset zip file
download.file(
  url = "https://api-proxy.edh-cde.dfo-mpo.gc.ca/catalogue/records/fed5f00f-7b17-4ac2-95d6-f1a73858dac0/attachments/Recreational_Boating_Data_Model.gdb.zip",
  destfile = zip_file,
  mode = "wb"
)

# Unzip the contents
dir.create(unzip_dir, showWarnings = FALSE)
unzip(zip_file, exdir = unzip_dir)

# List the file names you just downloaded 
boat_file = list.files(unzip_dir, full.names = TRUE)

# Turn the "survey effort" data into raster format
surveyRast = rast(boat_file, sub = "surveyeffort")

# Plot it just to see what it looks like
# It's gross! There are lots of values with no survey effort that make everything yucky to look at
plot(surveyRast)

# We need to fix up the data a bit more. First, unfortunately SpatRasters can't directly be plotted with ggplot
# So, we have to turn this into into a data frame
# But we can't do that directly, because the current projection is UTM, which means we'll lose lat/lon info
# So we first need to project it, so it's in the same CRS as the coast layer

# Use the project function to project to WGS 84 (i.e., the CRS of the coast file)
rast_proj = project(surveyRast, coast)

# You'll see this also fixes the issue of no data. Things get cleared up!
plot(rast_proj)

# Convert it to a dataframe. Keep the coordinate info
rast_df = as.data.frame(rast_proj, xy = TRUE) 

# Plot it with geom_tile and add the coast data as well
ggplot()+
  geom_tile(data = rast_df, aes(x=x, y=y, fill = Recreational_Boating_Data_Model)) +
  geom_sf(data = coast, fill = "lightgrey")+
  scale_fill_viridis_c(option = "magma")+
  coord_sf(
    xlim = c(bounds["xmin"], bounds["xmax"]),
    ylim = c(bounds["ymin"], bounds["ymax"])
  )+
  theme_bw()

```

## Making Leaflet Maps
``` {r, echo = TRUE, message= FALSE, warning= FALSE, error= FALSE, cache= FALSE, results='hide'}

# Convert raster to leaflet-compatible format
pal = colorNumeric(palette = "magma", domain = values(rast_proj), na.color = "transparent")

# Create leaflet map
leaflet() %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>%
  addRasterImage(rast_proj, colors = pal, opacity = 0.7, project = T) %>%
  addLegend(pal = pal, values = values(rast_proj),
            title = "Boating Effort",
            position = "bottomright")

```

## Searching for data with ckanr


################################################################################
### Using ckanr 

Our IT sometimes need to upgrade the server hardware or software, which can cause the URLs to the data to change (e.g. this part: "https://egisp.dfo-mpo.gc.ca/arcgis/..."), but as far as I know the Universally Unique Identifier (UUID) will persist. We can leverage the [CKAN](https://ckan.org/) API to make our script above more robust. CKAN is a data management system that allows you to publish, share, and find data. The `ckanr` package allows you to interact with CKAN from R. You can use the `ckanr_setup()` function to set up the connection to the Open Data portal (amongst others). 

### Finding data

While you can search for data on the [Open Data](https://search.open.canada.ca/data/) portal, it can be cumbersome to find the data you need. The [Open Science Data Portal](https://osdp-psdo.canada.ca/dp/en/search) (OSDP) is actually much better for finding data, searches the same database, and the UUID'd are consistent among the search portals. 

``` {r, echo = TRUE, message= FALSE, warning= FALSE, error= FALSE, cache= FALSE, results='hide'}

# If you don't want to search via the browser, you can use the `ckanr` package to search for data using the `package_search()` function. You can use the `q` argument to search for data. For example, you can search for "cumulative human impact" to find the data we will use in this example.
ckanr_setup("https://open.canada.ca/data")
search_results = package_search(q = "Pacific salmon")
search_results

# This shows how to find spatial data (e.g., shapefiles, geodatabases, etc.)
search_results = package_search(
  q = "Pacific salmon",
  fq = 'res_format:(SHP OR GDB OR GeoJSON OR KML OR CSV)' # CSV is sometimes spatial (lat/lon), sometimes not
)
```